---
title: "COOP"
author: "Brian"
date: "7/27/2022"
output:
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

# Disclaimer 

*This work is very preliminary as I get back into the coding swing of things. Data wrangling and figure generation will be done via R, but the rest of the project will be done using good ol' microsoft products. This is just an entry point into data crunching and should by no means be considered a final product.*

3/30/2023 update, corrected a minor error in methodology.

# Steamboat COOP

```{r libraries, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(snotelr)
library(riem)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(dataRetrieval)
library(lubridate)
library(sf)
library(ggthemes)
library(xts)
library(dygraphs)
library(scales)
library(openair)
library(plotly)
library(SciViews)
knitr::opts_chunk$set(message = F, 
                      warning = F,
                      cache = T)
knitr::opts_chunk$set(echo = TRUE)
library(trend)
library(nhdplusTools)
library(lfstat)
library(ggpubr)
library(kableExtra)

#Stats
library(forcats)
library(stringr)
library(trend)

# COOP/ Did not work
#library(rnoaa)
```

# Coop Data

NOAA/NWS Cooperative Observer Network

Two stations near steamboat, according to this [site](https://wrcc.dri.edu/inventory/coop05.html). 
057936 02 SBTC2 UNITED STATES CO ROUTT +7 STEAMBOAT SPRINGS 40 30 17 -106 51 58  6636
057942 02 SSPC2 UNITED STATES CO ROUTT +7 STEAMBOAT SPRINGS 1 W 40 29 00 -106 51 00 6700

Downloading from the [Iowa Environmental Mesonet](https://mesonet.agron.iastate.edu/request/coop/fe.phtml?network=COCLIMATE). NOAA package not working, maybe due to user error. 

```{r Read in Steamboat COOP, eval=FALSE, include=FALSE}
# coops_search(station_name = 057936,product = "air_temperature") # Handy R package did not work.

# At home:
coop_steamboat <- read.csv("C:/Users/13074/Documents/ESS580/thesis_project/research-COOP/data_raw/steamboat_nwscoop_1900_2022.csv", header = TRUE)

# From Campus:
#coop_steamboat <- read.csv("C:/Users/bsteen/OneDrive - Colostate/Documents/MS_Research/research-COOP/data_raw/steamboat_nwscoop_1900_2022.csv", header = TRUE)


```

### Data Cleaning

```{r data clean, eval=FALSE, include=FALSE}

str(coop_steamboat) # check the date, usually a character.

coop_steamboat$Date <- mdy(coop_steamboat$day)

coop_steamboat_clean <- coop_steamboat %>% # filter for the timeframe
  addWaterYear() %>%
  mutate(daymonth = format(as.Date(Date), "%d-%m")) %>% 
  group_by(waterYear)%>% 
  mutate(waterDay = (as.integer(difftime(Date, ymd(paste0(waterYear - 1 ,'-09-30')), units = "days")))) %>%   na.omit()

#For home:
write.csv(coop_steamboat_clean,"C:/Users/13074/Documents/ESS580/thesis_project/research-COOP/data_clean/coop_steamboat_clean.csv", row.names = FALSE)

# for campus:
#write.csv(coop_steamboat_clean,"C:/Users/bsteen/OneDrive - #Colostate/Documents/MS_Research/research-COOP/data_clean/coop_steamboat_clean.csv", row.names = FALSE)

```

```{r read in on campus}

#coop_steamboat_clean <- read.csv("C:/Users/bsteen/OneDrive - #Colostate/Documents/MS_Research/research-COOP/data_clean/coop_steamboat_clean.csv", header = TRUE)

```


### Figure check

```{r S COOP simple plot}

# from home
coop_steamboat_clean <- read.csv("C:/Users/13074/Documents/ESS580/thesis_project/research-COOP/data_clean/coop_steamboat_clean.csv", header = TRUE)

coop_steamboat_clean$Date <- ymd(coop_steamboat_clean$Date)

str(coop_steamboat_clean)

ggplot(coop_steamboat_clean, aes(x = Date, y = avg_T_c)) +
  geom_line() + #lwd = 2) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('Daily temperature (°C)') + 
  xlab('Date')

#Check for outliers....

#dygraph

COOP_temp_xts <- xts(coop_steamboat_clean$avg_T_c, order.by = coop_steamboat_clean$Date)

dygraph(COOP_temp_xts) %>%
  dyAxis("y", label = "Daily temperature (°C)") 

```

## Detrending Data 

```{r detrending data}
#SF figured out the yearly average by water year

#average water year temperature

COOP_yearly_wy_aver <- coop_steamboat_clean %>% 
  group_by(waterYear) %>% 
  mutate(aver_ann_temp = mean(avg_T_c))

```

```{r detrending data2}
#Average temperature by day for all water years:

COOP_daily_wy_aver <- COOP_yearly_wy_aver %>% 
  group_by(daymonth) %>% 
  mutate(aver_day_temp = mean(avg_T_c))

#average mean temperature by day for the period of record:

COOP_daily_wy_aver <- COOP_daily_wy_aver %>% 
  group_by(daymonth) %>% 
  mutate(all_ave_temp = mean(COOP_daily_wy_aver$aver_day_temp))

#str(daily_wy_aver)

```

```{r Figure all year average temp}
# try to show all years as means. 
COOP_daily_wy_aver2 <- COOP_daily_wy_aver %>% 
  #filter(waterYear == "1987" | waterYear == "2021") %>%
  group_by(waterDay) %>%
  mutate(date_temp = mean(avg_T_c))
  
COOP_daily_wy_aver2$date_temp <- signif(COOP_daily_wy_aver2$date_temp,3) #reduce the sig figs

ggplot(COOP_daily_wy_aver2, aes(x = waterDay, y = date_temp))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  #geom_smooth(method = "lm", se=FALSE) +
  ylab('Average Daily temperature (°C)') + 
  xlab('Day of water year')
```


```{r xts Figure all year average temp, eval=FALSE, include=FALSE}
COOP_temp_xts_2 <- xts(COOP_daily_wy_aver2$date_temp, order.by = as.Date(COOP_daily_wy_aver2$waterDay))

dygraph(COOP_temp_xts_2) %>%
  dyAxis("y", label = "(°C)") 

# **Day of year average temperature for the 1987-2021 period of record for Steamboat COOP site. *(x-axis is day of year, not date)* **
```

# Standard Deviation 

To figure out the standard deviation for each year, I want the "residual" for each daily value. 

The standard deviation will be the daily residual minus the mean of the residuals by water year, summed and squared, then divided by the number of observations minus one. The square root of the resulting value of which is thus the standard deviation for the water year. 

Determining residuals
```{r COOP residuals}
COOP_standard_dev <- COOP_daily_wy_aver %>% 
  group_by(waterYear) %>% 
  mutate(residual = (all_ave_temp-aver_ann_temp)+avg_T_c-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual)))

mean(COOP_standard_dev$residual)
```
*The mean of the residuals is close enough to zero*

Calculating standard deviation for the timeseries
```{r COOP sd}
COOP_standard_dev_all <- COOP_standard_dev %>% 
  group_by(waterYear) %>% 
  mutate(nmbr = n())

COOP_standard_dev_all <- COOP_standard_dev_all %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)

COOP_standard_dev_all %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')

ggplot(COOP_standard_dev_all, aes(x = waterYear, y = sd_2))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')
```



## Checking a random year from the timeseries. 
```{r sd 87}
COOP_standard_dev_87 <- COOP_standard_dev %>% 
  filter(waterYear == 1987) %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
           mutate(sd_1 = residual-resid_mean)

COOP_standard_dev_87 <- COOP_standard_dev_87 %>%
  group_by(waterYear) %>%
  mutate(sd_2 = (((sum((sd_1)^2))/((sum(tabulate(COOP_standard_dev_87$waterDay)))-1)))^(0.5))

head(COOP_standard_dev_87$sd_2, 1)
```
*Looks good.*

# Summer temperature standard deviation

```{r COOP summer}

COOP_standard_dev_all_summer <- COOP_standard_dev %>%
  filter(waterDay >= 244 & waterDay <= 335) %>% # this might be better off as daymonth rather than day of water year due to leap year
  group_by(waterYear) %>% 
  mutate(nmbr = n())

COOP_standard_dev_all_summer <- COOP_standard_dev_all_summer %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)

COOP_standard_dev_all_summer %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')

ggplot(COOP_standard_dev_all_summer, aes(x = waterYear, y = sd_2))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')

```

**Jun-Aug standard deviation for water years 1900-2021**

## Mann-Kendall & Sen’s Slope

Summer standard deviations.
```{r sd mk & ss summer}

sd_mk_summer <- mk.test(COOP_standard_dev_all_summer$sd_2)
print(sd_mk_summer)

sd_sens_summer <- sens.slope(COOP_standard_dev_all_summer$sd_2)
print(sd_sens_summer)

```

# Winter temperature standard deviation

```{r COOP winter}

COOP_standard_dev_all_winter <- COOP_standard_dev %>%
  filter(waterDay >= 32 & waterDay <= 182) %>% # this might be better off as daymonth rather than day of water year due to leap year
  group_by(waterYear) %>% 
  mutate(nmbr = n())

# Nope. This did some weird stuff with twice the observations.
# COOP_standard_dev_all_winter <- COOP_standard_dev %>%
#   filter(daymonth >= "01-11" & daymonth <= "31-03") %>%
#   group_by(waterYear) %>% 
#   mutate(nmbr = n())


COOP_standard_dev_all_winter <- COOP_standard_dev_all_winter %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)

COOP_standard_dev_all_winter %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')

ggplot(COOP_standard_dev_all_winter, aes(x = waterYear, y = sd_2))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')

```

**Nov-Mar standard deviation for water years 1900-2021**

## Mann-Kendall & Sen’s Slope

Winter standard deviations.
```{r sd mk & ss winter}

sd_mk_winter <- mk.test(COOP_standard_dev_all_winter$sd_2)
print(sd_mk_winter)

sd_sens_winter <- sens.slope(COOP_standard_dev_all_winter$sd_2)
print(sd_sens_winter)

```

Smaller time increments might be useful....

# 1987 - 2021

In comparison to the Tower SNOTEL site: 

```{r COOP to Tower}

# I did this wrong- need to compare the 1987- 2021 time series without the mean of the 1900 - 2021 time series influencing the residuals.  

#average water year temperature

COOP_yearly_wy_aver_87_21 <- coop_steamboat_clean %>%
  filter(waterYear >= 1987 & waterYear <= 2021) %>% 
  group_by(waterYear) %>% 
  mutate(aver_ann_temp = mean(avg_T_c))

#Average temperature by day for all water years:

COOP_daily_wy_aver_87_21 <- COOP_yearly_wy_aver_87_21 %>% 
  group_by(daymonth) %>% 
  mutate(aver_day_temp = mean(avg_T_c))

#average mean temperature by day for the period of record:

COOP_daily_wy_aver_87_21 <- COOP_daily_wy_aver_87_21 %>% 
  group_by(daymonth) %>% 
  mutate(all_ave_temp = mean(COOP_daily_wy_aver_87_21$aver_day_temp)) %>% 
  group_by(waterYear) %>% 
  mutate(nmbr = n())
  

COOP_standard_dev_87_21 <- COOP_daily_wy_aver_87_21 %>% 
  group_by(waterYear) %>% 
  mutate(residual = (all_ave_temp-aver_ann_temp)+avg_T_c-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual)))

COOP_standard_dev_87_21 <- COOP_standard_dev_87_21 %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)


COOP_standard_dev_87_21 %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')

ggplot(COOP_standard_dev_87_21, aes(x = waterYear, y = sd_2))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')

```

**SD for water years 1987 - 2021**

1987 - 2021  Mann-Kendall & Sen’s Slope

```{r COOP to Tower tests}

sd_mk_87_21 <- mk.test(COOP_standard_dev_87_21$sd_2)
print(sd_mk_87_21)

sd_sens_87_21 <- sens.slope(COOP_standard_dev_87_21$sd_2)
print(sd_sens_87_21)

```

## Summer and Winter 87-21

# Summer temperature standard deviation

```{r COOP 87-21 summer}

# using the 1987- 2021 data frame

COOP_standard_dev_all_87_21_summer <- COOP_daily_wy_aver_87_21 %>%
  filter(waterDay >= 244 & waterDay <= 335) %>% # this might be better off as daymonth rather than day of water year due to leap year
  group_by(waterYear) %>% 
  mutate(residual = (all_ave_temp-aver_ann_temp)+avg_T_c-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual))) %>% 
  group_by(waterYear) %>% 
  mutate(nmbr = n())

COOP_standard_dev_all_87_21_summer <- COOP_standard_dev_all_87_21_summer %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)

COOP_standard_dev_all_87_21_summer %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')

ggplot(COOP_standard_dev_all_87_21_summer, aes(x = waterYear, y = sd_2))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')

```

**Summer season SD for water years 1987 - 2021**

## Mann-Kendall & Sen’s Slope

Summer 87-21 standard deviations.
```{r COOP sd mk & ss 87-21 summer}

sd_mk_summer <- mk.test(COOP_standard_dev_all_87_21_summer$sd_2)
print(sd_mk_summer)

sd_sens_summer <- sens.slope(COOP_standard_dev_all_summer$sd_2)
print(sd_sens_summer)

```

# Winter temperature standard deviation

```{r COOP 87-21 winter}

# using the 1987- 2021 data frame

COOP_standard_dev_all_87_21_winter <- COOP_daily_wy_aver_87_21 %>%
  filter(waterDay >= 32 & waterDay <= 182) %>% # this might be better off as daymonth rather than day of water year due to leap year
  group_by(waterYear) %>% 
  mutate(residual = (all_ave_temp-aver_ann_temp)+avg_T_c-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual))) %>% 
  group_by(waterYear) %>% 
  mutate(nmbr = n())

COOP_standard_dev_all_87_21_winter <- COOP_standard_dev_all_87_21_winter %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)

COOP_standard_dev_all_87_21_winter %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')

ggplot(COOP_standard_dev_all_87_21_winter, aes(x = waterYear, y = sd_2))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')

```

**Winter season SD for water years 1987 - 2021**

## Mann-Kendall & Sen’s Slope

1987 - 2021 winter Mann-Kendall & Sen’s Slope

```{r COOP sd mk & ss 87-21 winter}

sd_mk_87_21_winter <- mk.test(COOP_standard_dev_all_87_21_winter$sd_2)
print(sd_mk_87_21_winter)

sd_sens_87_21_winter <- sens.slope(COOP_standard_dev_all_87_21_winter$sd_2)
print(sd_sens_87_21_winter)

```

Neither the 1987 - 2021 as a whole, nor the seasonal time series demonstrate trends that are statistically significant. 

# 1900-1915 minimum and maximum temperatures

Jan 7, 1913 had the lowest recorded temperature. This just looks at the max and min temperatures for that early part of the record. 
```{r 1913 min}

early_coop <- coop_steamboat_clean %>% 
  filter(waterYear >= 1900 & waterYear <= 1915)

COOP_temp_min_xts <- xts(early_coop$min_T_c, order.by = early_coop$Date)

dygraph(COOP_temp_min_xts) %>%
  dyAxis("y", label = "Daily min temperature (°C)") 


```

*Daily minimum temperatures for water years 1900-1915*

```{r 1913 max}

COOP_temp_max_xts <- xts(early_coop$max_T_c, order.by = early_coop$Date)

dygraph(COOP_temp_max_xts) %>%
  dyAxis("y", label = "Daily max temperature (°C)") 


```

*Daily minimum temperatures for water years 1900-1915*


# **By groups of decades**

## 1900-1930 

```{r COOP 1900-1930}

#average water year temperature

COOP_yearly_wy_aver_00_30 <- coop_steamboat_clean %>%
  filter(waterYear >= 1900 & waterYear <= 1930) %>% 
  group_by(waterYear) %>% 
  mutate(aver_ann_temp = mean(avg_T_c))

#Average temperature by day for all water years:

COOP_daily_wy_aver_00_30 <- COOP_yearly_wy_aver_00_30 %>% 
  group_by(daymonth) %>% 
  mutate(aver_day_temp = mean(avg_T_c))

#average mean temperature by day for the period of record:

COOP_daily_wy_aver_00_30 <- COOP_daily_wy_aver_00_30 %>% 
  group_by(daymonth) %>% 
  mutate(all_ave_temp = mean(COOP_daily_wy_aver_00_30$aver_day_temp)) %>% 
  group_by(waterYear) %>% 
  mutate(nmbr = n())
  

COOP_standard_dev_00_30 <- COOP_daily_wy_aver_00_30 %>% 
  group_by(waterYear) %>% 
  mutate(residual = (all_ave_temp-aver_ann_temp)+avg_T_c-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual)))

COOP_standard_dev_00_30 <- COOP_standard_dev_00_30 %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)


COOP_standard_dev_00_30 %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')

ggplot(COOP_standard_dev_00_30, aes(x = waterYear, y = sd_2))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')

```

**SD for water years 1900 - 1930**

1900 - 1930  Mann-Kendall & Sen’s Slope

```{r COOP 00-30 tests}

sd_mk_00_30 <- mk.test(COOP_standard_dev_00_30$sd_2)
print(sd_mk_00_30)

sd_sens_00_30 <- sens.slope(COOP_standard_dev_00_30$sd_2)
print(sd_sens_00_30)

```

### Summer and Winter 00-30

Summer temperature standard deviation

```{r COOP 00-30 summer}

# using the 1900- 1930 data frame

COOP_standard_dev_all_00_30_summer <- COOP_daily_wy_aver_00_30 %>%
  filter(waterDay >= 244 & waterDay <= 335) %>% # this might be better off as daymonth rather than day of water year due to leap year
  group_by(waterYear) %>% 
  mutate(residual = (all_ave_temp-aver_ann_temp)+avg_T_c-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual))) %>% 
  group_by(waterYear) %>% 
  mutate(nmbr = n())

COOP_standard_dev_all_00_30_summer <- COOP_standard_dev_all_00_30_summer %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)

COOP_standard_dev_all_00_30_summer %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')

ggplot(COOP_standard_dev_all_00_30_summer, aes(x = waterYear, y = sd_2))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')

```

**Summer season SD for water years 1900 - 1930**

Mann-Kendall & Sen’s Slope

Summer 00-30 standard deviations.
```{r COOP sd mk & ss 00-30 summer}

sd_mk_00_30_summer <- mk.test(COOP_standard_dev_all_00_30_summer$sd_2)
print(sd_mk_00_30_summer)

sd_sens_00_30_summer <- sens.slope(COOP_standard_dev_all_00_30_summer$sd_2)
print(sd_sens_00_30_summer)

```

 Winter temperature standard deviation

```{r COOP 00-30 winter}

# using the 00-30 data frame

COOP_standard_dev_all_00_30_winter <- COOP_daily_wy_aver_00_30 %>%
  filter(waterDay >= 32 & waterDay <= 182) %>% # this might be better off as daymonth rather than day of water year due to leap year
  group_by(waterYear) %>% 
  mutate(residual = (all_ave_temp-aver_ann_temp)+avg_T_c-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual))) %>% 
  group_by(waterYear) %>% 
  mutate(nmbr = n())

COOP_standard_dev_all_00_30_winter <- COOP_standard_dev_all_00_30_winter %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)

COOP_standard_dev_all_00_30_winter %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')

ggplot(COOP_standard_dev_all_00_30_winter, aes(x = waterYear, y = sd_2))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')

```

**Winter season SD for water years 1900 - 1930**

 Mann-Kendall & Sen’s Slope

1900 - 1930 winter Mann-Kendall & Sen’s Slope

```{r COOP sd mk & ss 00-30 winter}

sd_mk_00_30_winter <- mk.test(COOP_standard_dev_all_00_30_winter$sd_2)
print(sd_mk_00_30_winter)

sd_sens_00_30_winter <- sens.slope(COOP_standard_dev_all_00_30_winter$sd_2)
print(sd_sens_00_30_winter)

```


## 1931-1960 

```{r COOP 1931-1960}

#average water year temperature

COOP_yearly_wy_aver_31_60 <- coop_steamboat_clean %>%
  filter(waterYear >= 1931 & waterYear <= 1960) %>% 
  group_by(waterYear) %>% 
  mutate(aver_ann_temp = mean(avg_T_c))

#Average temperature by day for all water years:

COOP_daily_wy_aver_31_60 <- COOP_yearly_wy_aver_31_60 %>% 
  group_by(daymonth) %>% 
  mutate(aver_day_temp = mean(avg_T_c))

#average mean temperature by day for the period of record:

COOP_daily_wy_aver_31_60 <- COOP_daily_wy_aver_31_60 %>% 
  group_by(daymonth) %>% 
  mutate(all_ave_temp = mean(COOP_daily_wy_aver_31_60$aver_day_temp)) %>% 
  group_by(waterYear) %>% 
  mutate(nmbr = n())
  

COOP_standard_dev_31_60 <- COOP_daily_wy_aver_31_60 %>% 
  group_by(waterYear) %>% 
  mutate(residual = (all_ave_temp-aver_ann_temp)+avg_T_c-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual)))

COOP_standard_dev_31_60 <- COOP_standard_dev_31_60 %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)


COOP_standard_dev_31_60 %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')

ggplot(COOP_standard_dev_31_60, aes(x = waterYear, y = sd_2))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')

```

**SD for water years 1931 - 1960**

1931 - 1960  Mann-Kendall & Sen’s Slope

```{r COOP 31-60 tests}

sd_mk_31_60 <- mk.test(COOP_standard_dev_31_60$sd_2)
print(sd_mk_31_60)

sd_sens_31_60 <- sens.slope(COOP_standard_dev_31_60$sd_2)
print(sd_sens_31_60)

```

### Summer and Winter 31-60

 Summer temperature standard deviation

```{r COOP 31-60 summer}

# using the 1931- 1960 data frame

COOP_standard_dev_all_31_60_summer <- COOP_daily_wy_aver_31_60 %>%
  filter(waterDay >= 244 & waterDay <= 335) %>% # this might be better off as daymonth rather than day of water year due to leap year
  group_by(waterYear) %>% 
  mutate(residual = (all_ave_temp-aver_ann_temp)+avg_T_c-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual))) %>% 
  group_by(waterYear) %>% 
  mutate(nmbr = n())

COOP_standard_dev_all_31_60_summer <- COOP_standard_dev_all_31_60_summer %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)

COOP_standard_dev_all_31_60_summer %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')

ggplot(COOP_standard_dev_all_31_60_summer, aes(x = waterYear, y = sd_2))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')

```

**Summer season SD for water years 1931 - 1960**

Mann-Kendall & Sen’s Slope

Summer 31-60 standard deviations.
```{r COOP sd mk & ss 31-60 summer}

sd_mk_31_60_summer <- mk.test(COOP_standard_dev_all_31_60_summer$sd_2)
print(sd_mk_31_60_summer)

sd_sens_31_60_summer <- sens.slope(COOP_standard_dev_all_31_60_summer$sd_2)
print(sd_sens_31_60_summer)

```

Winter temperature standard deviation

```{r COOP 31-60 winter}

# using the 31-60 data frame

COOP_standard_dev_all_31_60_winter <- COOP_daily_wy_aver_31_60 %>%
  filter(waterDay >= 32 & waterDay <= 182) %>% # this might be better off as daymonth rather than day of water year due to leap year
  group_by(waterYear) %>% 
  mutate(residual = (all_ave_temp-aver_ann_temp)+avg_T_c-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual))) %>% 
  group_by(waterYear) %>% 
  mutate(nmbr = n())

COOP_standard_dev_all_31_60_winter <- COOP_standard_dev_all_31_60_winter %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)

COOP_standard_dev_all_31_60_winter %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')

ggplot(COOP_standard_dev_all_31_60_winter, aes(x = waterYear, y = sd_2))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')

```

**Winter season SD for water years 1931 - 1960**

Mann-Kendall & Sen’s Slope

1931 - 1960 winter Mann-Kendall & Sen’s Slope

```{r COOP sd mk & ss 31-60 winter}

sd_mk_31_60_winter <- mk.test(COOP_standard_dev_all_31_60_winter$sd_2)
print(sd_mk_31_60_winter)

sd_sens_31_60_winter <- sens.slope(COOP_standard_dev_all_31_60_winter$sd_2)
print(sd_sens_31_60_winter)

```


## 1961-1990 


```{r COOP 1961-1990}

#average water year temperature

COOP_yearly_wy_aver_61_90 <- coop_steamboat_clean %>%
  filter(waterYear >= 1961 & waterYear <= 1990) %>% 
  group_by(waterYear) %>% 
  mutate(aver_ann_temp = mean(avg_T_c))

#Average temperature by day for all water years:

COOP_daily_wy_aver_61_90 <- COOP_yearly_wy_aver_61_90 %>% 
  group_by(daymonth) %>% 
  mutate(aver_day_temp = mean(avg_T_c))

#average mean temperature by day for the period of record:

COOP_daily_wy_aver_61_90 <- COOP_daily_wy_aver_61_90 %>% 
  group_by(daymonth) %>% 
  mutate(all_ave_temp = mean(COOP_daily_wy_aver_61_90$aver_day_temp)) %>% 
  group_by(waterYear) %>% 
  mutate(nmbr = n())
  

COOP_standard_dev_61_90 <- COOP_daily_wy_aver_61_90 %>% 
  group_by(waterYear) %>% 
  mutate(residual = (all_ave_temp-aver_ann_temp)+avg_T_c-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual)))

COOP_standard_dev_61_90 <- COOP_standard_dev_61_90 %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)


COOP_standard_dev_61_90 %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')

ggplot(COOP_standard_dev_61_90, aes(x = waterYear, y = sd_2))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')

```

**SD for water years 1961 - 1990**

1961 - 1990  Mann-Kendall & Sen’s Slope

```{r COOP 61-90 tests}

sd_mk_61_90 <- mk.test(COOP_standard_dev_61_90$sd_2)
print(sd_mk_61_90)

sd_sens_61_90 <- sens.slope(COOP_standard_dev_61_90$sd_2)
print(sd_sens_61_90)

```

### Summer and Winter 61-90

Summer temperature standard deviation

```{r COOP 61-90 summer}

# using the 1961- 1990 data frame

COOP_standard_dev_all_61_90_summer <- COOP_daily_wy_aver_61_90 %>%
  filter(waterDay >= 244 & waterDay <= 335) %>% # this might be better off as daymonth rather than day of water year due to leap year
  group_by(waterYear) %>% 
  mutate(residual = (all_ave_temp-aver_ann_temp)+avg_T_c-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual))) %>% 
  group_by(waterYear) %>% 
  mutate(nmbr = n())

COOP_standard_dev_all_61_90_summer <- COOP_standard_dev_all_61_90_summer %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)

COOP_standard_dev_all_61_90_summer %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')

ggplot(COOP_standard_dev_all_61_90_summer, aes(x = waterYear, y = sd_2))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')

```

**Summer season SD for water years 1961 - 1990**

Mann-Kendall & Sen’s Slope

Summer 61-90 standard deviations.
```{r COOP sd mk & ss 61-90 summer}

sd_mk_61_90_summer <- mk.test(COOP_standard_dev_all_61_90_summer$sd_2)
print(sd_mk_61_90_summer)

sd_sens_61_90_summer <- sens.slope(COOP_standard_dev_all_61_90_summer$sd_2)
print(sd_sens_61_90_summer)

```

Winter temperature standard deviation

```{r COOP 61-90 winter}

# using the 61-90 data frame

COOP_standard_dev_all_61_90_winter <- COOP_daily_wy_aver_61_90 %>%
  filter(waterDay >= 32 & waterDay <= 182) %>% # this might be better off as daymonth rather than day of water year due to leap year
  group_by(waterYear) %>% 
  mutate(residual = (all_ave_temp-aver_ann_temp)+avg_T_c-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual))) %>% 
  group_by(waterYear) %>% 
  mutate(nmbr = n())

COOP_standard_dev_all_61_90_winter <- COOP_standard_dev_all_61_90_winter %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)

COOP_standard_dev_all_61_90_winter %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')

ggplot(COOP_standard_dev_all_61_90_winter, aes(x = waterYear, y = sd_2))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')

```

**Winter season SD for water years 1961 - 1990**

Mann-Kendall & Sen’s Slope

1961 - 1990 winter Mann-Kendall & Sen’s Slope

```{r COOP sd mk & ss 61-90 winter}

sd_mk_61_90_winter <- mk.test(COOP_standard_dev_all_61_90_winter$sd_2)
print(sd_mk_61_90_winter)

sd_sens_61_90_winter <- sens.slope(COOP_standard_dev_all_61_90_winter$sd_2)
print(sd_sens_61_90_winter)

```

## 1991-2020

```{r COOP 1991-2020}

#average water year temperature

COOP_yearly_wy_aver_91_20 <- coop_steamboat_clean %>%
  filter(waterYear >= 1991 & waterYear <= 2020) %>% 
  group_by(waterYear) %>% 
  mutate(aver_ann_temp = mean(avg_T_c))

#Average temperature by day for all water years:

COOP_daily_wy_aver_91_20 <- COOP_yearly_wy_aver_91_20 %>% 
  group_by(daymonth) %>% 
  mutate(aver_day_temp = mean(avg_T_c))

#average mean temperature by day for the period of record:

COOP_daily_wy_aver_91_20 <- COOP_daily_wy_aver_91_20 %>% 
  group_by(daymonth) %>% 
  mutate(all_ave_temp = mean(COOP_daily_wy_aver_91_20$aver_day_temp)) %>% 
  group_by(waterYear) %>% 
  mutate(nmbr = n())
  

COOP_standard_dev_91_20 <- COOP_daily_wy_aver_61_90 %>% 
  group_by(waterYear) %>% 
  mutate(residual = (all_ave_temp-aver_ann_temp)+avg_T_c-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual)))

COOP_standard_dev_91_20 <- COOP_standard_dev_91_20 %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)


COOP_standard_dev_91_20 %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')

ggplot(COOP_standard_dev_91_20, aes(x = waterYear, y = sd_2))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')

```

**SD for water years 1991 - 2020**

1991 - 2020  Mann-Kendall & Sen’s Slope

```{r COOP 91-20 tests}

sd_mk_91_20 <- mk.test(COOP_standard_dev_91_20$sd_2)
print(sd_mk_91_20)

sd_sens_91_20 <- sens.slope(COOP_standard_dev_91_20$sd_2)
print(sd_sens_91_20)

```

### Summer and Winter 91-20

Summer temperature standard deviation

```{r COOP 91-20 summer}

# using the 1991- 2020 data frame

COOP_standard_dev_all_91_20_summer <- COOP_daily_wy_aver_91_20 %>%
  filter(waterDay >= 244 & waterDay <= 335) %>% # this might be better off as daymonth rather than day of water year due to leap year
  group_by(waterYear) %>% 
  mutate(residual = (all_ave_temp-aver_ann_temp)+avg_T_c-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual))) %>% 
  group_by(waterYear) %>% 
  mutate(nmbr = n())

COOP_standard_dev_all_91_20_summer <- COOP_standard_dev_all_91_20_summer %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)

COOP_standard_dev_all_91_20_summer %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')

ggplot(COOP_standard_dev_all_91_20_summer, aes(x = waterYear, y = sd_2))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')

```

**Summer season SD for water years 1991 - 2020**

Mann-Kendall & Sen’s Slope

Summer 91-20 standard deviations.
```{r COOP sd mk & ss 91-20 summer}

sd_mk_91_20_summer <- mk.test(COOP_standard_dev_all_91_20_summer$sd_2)
print(sd_mk_91_20_summer)

sd_sens_91_20_summer <- sens.slope(COOP_standard_dev_all_91_20_summer$sd_2)
print(sd_sens_91_20_summer)

```

Winter temperature standard deviation

```{r COOP 91-20 winter}

# using the 91-20 data frame

COOP_standard_dev_all_91_20_winter <- COOP_daily_wy_aver_91_20 %>%
  filter(waterDay >= 32 & waterDay <= 182) %>% # this might be better off as daymonth rather than day of water year due to leap year
  group_by(waterYear) %>% 
  mutate(residual = (all_ave_temp-aver_ann_temp)+avg_T_c-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual))) %>% 
  group_by(waterYear) %>% 
  mutate(nmbr = n())

COOP_standard_dev_all_91_20_winter <- COOP_standard_dev_all_91_20_winter %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)

COOP_standard_dev_all_91_20_winter %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')

ggplot(COOP_standard_dev_all_91_20_winter, aes(x = waterYear, y = sd_2))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')

```

**Winter season SD for water years 1991 - 2020**

Mann-Kendall & Sen’s Slope

1991 - 2020 winter Mann-Kendall & Sen’s Slope

```{r COOP sd mk & ss 91-20 winter}

sd_mk_91_20_winter <- mk.test(COOP_standard_dev_all_91_20_winter$sd_2)
print(sd_mk_91_20_winter)

sd_sens_91_20_winter <- sens.slope(COOP_standard_dev_all_91_20_winter$sd_2)
print(sd_sens_91_20_winter)

```

# Comparing Climate Normal means

(from 356:)

## 1991-2022

```{r 1991-2022 CN detrend}
COOP_yearly_wy_aver_1991_2022_CN <- coop_steamboat_clean %>%
  filter(waterYear >= 1991 & waterYear <= 2022) %>% 
  group_by(waterYear) %>% 
  mutate(aver_ann_temp = mean(avg_T_c))

#Average temperature by day for all water years:

COOP_daily_wy_aver_1991_2022_CN <- COOP_yearly_wy_aver_1991_2022_CN %>% 
  group_by(daymonth) %>% 
  mutate(aver_day_temp = mean(avg_T_c))

#average mean temperature by day for the period of record:

COOP_daily_wy_aver_1991_2022_CN <- COOP_daily_wy_aver_1991_2022_CN %>% 
  group_by(daymonth) %>% 
  mutate(all_ave_temp = mean(COOP_daily_wy_aver_1991_2022_CN$aver_day_temp)) %>% 
  group_by(waterYear) %>% 
  mutate(nmbr = n())


```


```{r CN 1991-2022 means figure}
COOP_daily_wy_aver_1991_2022_CN2 <- COOP_daily_wy_aver_1991_2022_CN %>% 
  #filter(waterYear == "1987" | waterYear == "2021") %>%
  group_by(waterDay) %>%
  mutate(date_temp = mean(avg_T_c))%>% 
  select(waterDay, date_temp) %>% 
  distinct(waterDay, .keep_all = TRUE)
  
COOP_daily_wy_aver_1991_2022_CN2$date_temp <- signif(COOP_daily_wy_aver_1991_2022_CN2$date_temp,3) #reduce the sig figs

ggplot(COOP_daily_wy_aver_1991_2022_CN2, aes(x = waterDay, y = date_temp))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  theme_few()

```

## 1961-1990

```{r 1961-1990 CN detrend}
COOP_yearly_wy_aver_1961_1990_CN <- coop_steamboat_clean %>%
  filter(waterYear >= 1961 & waterYear <= 1990) %>% 
  group_by(waterYear) %>% 
  mutate(aver_ann_temp = mean(avg_T_c))

#Average temperature by day for all water years:

COOP_daily_wy_aver_1961_1990_CN <- COOP_yearly_wy_aver_1961_1990_CN %>% 
  group_by(daymonth) %>% 
  mutate(aver_day_temp = mean(avg_T_c))

#average mean temperature by day for the period of record:

COOP_daily_wy_aver_1961_1990_CN <- COOP_daily_wy_aver_1961_1990_CN %>% 
  group_by(daymonth) %>% 
  mutate(all_ave_temp = mean(COOP_daily_wy_aver_1961_1990_CN$aver_day_temp)) %>% 
  group_by(waterYear) %>% 
  mutate(nmbr = n())


```


```{r CN 1961-1990 means figure}
COOP_daily_wy_aver_1961_1990_CN2 <- COOP_daily_wy_aver_1961_1990_CN %>% 
  #filter(waterYear == "1987" | waterYear == "2021") %>%
  group_by(waterDay) %>%
  mutate(date_temp = mean(avg_T_c))%>% 
  select(waterDay, date_temp) %>% 
  distinct(waterDay, .keep_all = TRUE)
  
COOP_daily_wy_aver_1961_1990_CN2$date_temp <- signif(COOP_daily_wy_aver_1961_1990_CN2$date_temp,3) #reduce the sig figs

ggplot(COOP_daily_wy_aver_1961_1990_CN2, aes(x = waterDay, y = date_temp))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  theme_few()

```

## 1931-1960

```{r 1931-1960 CN detrend}
COOP_yearly_wy_aver_1931_1960_CN <- coop_steamboat_clean %>%
  filter(waterYear >= 1931 & waterYear <= 1960) %>% 
  group_by(waterYear) %>% 
  mutate(aver_ann_temp = mean(avg_T_c))

#Average temperature by day for all water years:

COOP_daily_wy_aver_1931_1960_CN <- COOP_yearly_wy_aver_1931_1960_CN %>% 
  group_by(daymonth) %>% 
  mutate(aver_day_temp = mean(avg_T_c))

#average mean temperature by day for the period of record:

COOP_daily_wy_aver_1931_1960_CN <- COOP_daily_wy_aver_1931_1960_CN %>% 
  group_by(daymonth) %>% 
  mutate(all_ave_temp = mean(COOP_daily_wy_aver_1931_1960_CN$aver_day_temp)) %>% 
  group_by(waterYear) %>% 
  mutate(nmbr = n())


```


```{r CN 1931-1960 means figure}
COOP_daily_wy_aver_1931_1960_CN2 <- COOP_daily_wy_aver_1931_1960_CN %>% 
  #filter(waterYear == "1987" | waterYear == "2021") %>%
  group_by(waterDay) %>%
  mutate(date_temp = mean(avg_T_c))%>% 
  select(waterDay, date_temp)  %>% 
  distinct(waterDay, .keep_all = TRUE)
  
COOP_daily_wy_aver_1931_1960_CN2$date_temp <- signif(COOP_daily_wy_aver_1931_1960_CN2$date_temp,3) #reduce the sig figs

ggplot(COOP_daily_wy_aver_1931_1960_CN2, aes(x = waterDay, y = date_temp))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  theme_few()

```

## 1900-1930

```{r 1901-1930 CN detrend}
COOP_yearly_wy_aver_1901_1930_CN <- coop_steamboat_clean %>%
  filter(waterYear >= 1901 & waterYear <= 1930) %>% 
  group_by(waterYear) %>% 
  mutate(aver_ann_temp = mean(avg_T_c))

#Average temperature by day for all water years:

COOP_daily_wy_aver_1901_1930_CN <- COOP_yearly_wy_aver_1901_1930_CN %>% 
  group_by(daymonth) %>% 
  mutate(aver_day_temp = mean(avg_T_c))

#average mean temperature by day for the period of record:

COOP_daily_wy_aver_1901_1930_CN <- COOP_daily_wy_aver_1901_1930_CN %>% 
  group_by(daymonth) %>% 
  mutate(all_ave_temp = mean(COOP_daily_wy_aver_1901_1930_CN$aver_day_temp)) %>% 
  group_by(waterYear) %>% 
  mutate(nmbr = n())


```


```{r CN 1901-1930 means figure}
COOP_daily_wy_aver_1901_1930_CN2 <- COOP_daily_wy_aver_1901_1930_CN %>% 
  #filter(waterYear == "1987" | waterYear == "2021") %>%
  group_by(waterDay) %>%
  mutate(date_temp = mean(avg_T_c))%>% 
  select(waterDay, date_temp) %>% 
  distinct(waterDay, .keep_all = TRUE)


COOP_daily_wy_aver_1901_1930_CN2$date_temp <- signif(COOP_daily_wy_aver_1901_1930_CN2$date_temp,3) #reduce the sig figs

ggplot(COOP_daily_wy_aver_1901_1930_CN2, aes(x = waterDay, y = date_temp))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  theme_few()

```

```{r}

climate_norm <- left_join(COOP_daily_wy_aver_1991_2022_CN2, COOP_daily_wy_aver_1961_1990_CN2, by= 'waterDay') 

climate_norm <- left_join(climate_norm, COOP_daily_wy_aver_1931_1960_CN2, by= 'waterDay')

climate_norm <- left_join(climate_norm, COOP_daily_wy_aver_1901_1930_CN2, by= 'waterDay')

ggplot(climate_norm, aes(x=waterDay)) +  
         #scale_size_manual(values=c("FALSE"=5,"TRUE"=8)) +
  #scale_color_manual(values=c("FALSE"='#CC0000',"TRUE"='black')) +
geom_line(aes(y = date_temp.x), size =1, color = "orange")+
  geom_line(aes(y = date_temp.y), size =1, color = "darkgreen")+
  geom_line(aes(y = date_temp.x.x), size =1, color = "purple")+
  geom_line(aes(y = date_temp.y.y), size =1, color= "red")+
  theme_few()+
  scale_colour_manual("", breaks=c("date_temp.x","date_temp.y","date_temp.x.x", "date_temp.y.y"), values=c("date_temp.x" = "orange", "date_temp.y" = "darkgreen", "date_temp.x.x" = "purple", "date_temp.y.y" = "red"))+
  xlab("Day of Year")+
  ylab("Average temperature")



    
```

